\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-07-13 paper/paper.tex
Add missing $ to table
Streamline intro/motivation section

In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions displaying early failures due to "infant mortality" and wearout. By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures. By producing samples from the full posterior distribution, estimation of arbitrary functionals is straightforward.

We illustrate this method on hard drive reliability data from a cloud-based computing company.  The data contains both left-truncation and right-censoring.  Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous.

\subsection{Motivation}

Failure distributions of high-reliability consumer products can be difficult to assess.  Even when field data are available, information may be limited for a variety of reasons.  Certain brands may have few products in operation resulting in small sample sizes.  Other products could have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.  In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process. After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability.   While simple nonparametric methods are available, such as the Kaplan-Meier estimator, these are not suitable for forecasting.  They also do not allow for a straight forward statistical comparisons of different brands in order to select the most reliable product.

For real reliability data, often no failure analysis is available allowing objective discernment about when a failure is due to infant mortality. This presents a dilemma for analysts: the data may indicate more than one failure mode, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. Where this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

While the GLFP model provides flexibility and accords with a macro-level theory of product failure, it requires a lot of data to fit because of its complexity. Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations, imposing "soft" constraints on model parameters. This enables estimation for all product groups.  Populations with lots of data will be precisely estimated.  For products with limited data the posterior distribution of time-to-failure will contain greater uncertainty, while being shrunk toward a "pooled" model. Finally, in order to compare the reliability across brands, the Bayesian approach is appealing as samples obtained from the posterior distribution via Markov Chain Monte Carlo allow for easy numerical integration for computing  various quantities of interest. \\

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & 2.5\% & 50\% & 97.5\% \\ 
  \hline
\pi & 0.03 & 0.05 & 0.10 \\ 
 \beta_1 & 0.47 & 1.13 & 1.72 \\ 
  \beta_2 & 4.47 & 4.70 & 4.95 \\ 
  \mu_1 & 7.46 & 8.06 & 8.76 \\ 
  \mu_2 & 10.12 & 10.13 & 10.14 \\ 
   \hline
\end{tabular}
\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive Model 14.}
\label{table:1}
\end{table}

>>>>

% In this paper we present a method for joint estimation of multiple lifetime distributions based on the GLFP model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions displaying early failures due to "infant mortality" and wearout. By borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures. By producing samples from the full posterior distribution, estimation of arbitrary functionals is straightforward.
% 
% We illustrate this method on hard drive reliability data from a cloud-based computing company.  The data contains both left-truncation and right-censoring.  Also, the information in the data is unevenly distributed across the populations of interest making a hierarchical approach especially advantageous. 
% 
% \subsection{Motivation}
% 
% Failure distributions of high-reliability consumer products can be difficult to assess.  Even when field data are available, information may be limited for a variety of reasons.  Certain brands may have few products in operation resulting in small sample sizes.  Other products could have many observed units in operation but few observed failures or limited time under observation, which makes fitting a parametric lifetime model problematic.  In this paper, we propose the use of Bayesian hierarchical modeling to borrow information across products, thereby improving inferences on lifetime distributions where information is limited by small sample sizes, left-truncation, and right censoring.

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring in early life ("infant mortality") due to defects from the manufacturing process. After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability. This highlights the importance of choosing a sufficiently flexible model.

Nonparametric methods, which provide such flexibility are well-known, Kaplan-Meier for example. However, truncation, censoring and finite samples limit the scope of inference for these methods. Parametric models do not have these limitations. Truncated and/or censored data can be accomodated simply by adjusting the likelihood. Additionally, because they assume a parsimonious representation of the data-generating process, they provide simpler, more intuitive means by which to compare populations of interest, borrow strength and incorporate prior information.

Often no failure analysis is available allowing objective discernment about when a failure is due to infant mortality. This presents a dilemma for analysts: the data may indicate more than one failure mode, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. Where this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

The GLFP model has a meaningful parameterization while accomodating lifetime data with infant mortality and wearout but it can require a lot of data to fit because of this complexity. Bayesian methods offer a means of reducing the amount of data required to fit the model by leveraging multiple sources of information. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes "soft" constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained by  Markov Chain Monte Carlo. \\

@@

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & 2.5\% & 50\% & 97.5\% \\ 
  \hline
$\pi$ & 0.03 & 0.05 & 0.10 \\ 
 $\beta_1$ & 0.47 & 1.13 & 1.72 \\ 
  $\beta_2$ & 4.47 & 4.70 & 4.95 \\ 
  $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
  $\mu_2$ & 10.12 & 10.13 & 10.14 \\ 
   \hline
\end{tabular}
\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive Model 14.}
\label{table:1}
\end{table}
